How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-15 tse-paper/content.tex
improve chunk retrieval techniques section

% \subsubsection{Build Logs}
% often focussed on specific logs, sometimes mentioned because of effort
% developing a parser other times also because other tools used in
% further research are limiting the scope (reference paper with PIT
% tool).
% many focus on maven because of it's prevalence also in respect
% to other tools much research done on travis ci logs b/c of their
% abundance and open availability
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
% TODO: rework! extend / relate to further techniques
% (should follow from discussion)
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).
We use these techniques to extract text chunks from build logs that
represent a specific, targeted information.

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
In this article, we evaluate different techniques to automatically
retrieve pieces of information that appear literally in the build
logs, i.e., techniques that do not aggreagte, combine, or deducted
information.
We call such pieces of information from build logs
\emph{chunks}, and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

@@

\begin{table*}[]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}XlXlXX@{}}
\toprule
Name			     & Acronym & Identification Technique
& Granularity & Configuration & Source		  \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
Diff Approach		     & ---     & Line not present in successful
log, information retrieval  & Line	  & Logs from failing and
successful builds      \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table*}

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
The area of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this apprach, we implemented a chunk retrieval technique
for build logs.
Our implementation is based and highly influenced by the
\emph{Microsoft PROSE library}~\cite{prose2019webpage}.
The PROSE library is powered by the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.
% % It enables users to create programs without a priori programming
% % knowledge~\cite{mayer2015user}.
% In the context of text extraction
% % through regular expressions, Programming by Example relieves the
% % developer from having to understand the whole document structure to
% % solve a single extraction task~\cite{le2014flashextract:}.
% In this
% % work, we refer to our interpretation of Programming by Example as
% % \emph{PBE}\@.
% We investigate the suitability of PBE to retrieve
% % information chunks from build logs.
% In the following, we explain the
% % configuration and application of PBE to chunk retrieval from CI
% % build logs.

% TODO later: we should make sure to at least mention PROSE
% _somewhere_, it's crucial for why PBE is how it is
\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example, we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
an array of substrings of the log file text, representing the
substrings that should be retrieved by the synthesized program when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
\Cref{lst:prose-program-simplified} presents a simplified version
of a program learned from the examples in \Cref{lst:chunk-example}.
% % A program is consistent with an example if it returns the defined
% % output when executed on the defined
% % input~\cite{mitchell1982generalization}.
% PBE reports an error back to the user if it could not synthesize a
% consistent program.
% % The program synthesis builds on the FlashExtract DSL, which in turn
% % uses the FlashMeta algorithm.
% Both are described in
% % Section~\ref{sec:rw-prose}.

@@

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
We investigate a chunk retrieval technique based on the common
and simple Vector Space Model~\cite{schutze2008introduction}.
The aim is to retrieve those lines from a build log which are most
similar to examples given by the user.
% Text Similarity approaches are used to filter unstructured textual
% software artifacts.
% One common technique
% %
%artifacts.
% % One common and simple technique is the Vector Space Model.
% We
% % investigate when text similarity is a suitable technique to retrieve
% % information chunks from build logs.
% In the following we will explain
% % the concept of how we apply text similarity to information retrieval
% % from CI build logs, which we refer to as \emph{CTS}\@.

\subsubsection{Configuration}
To configure chunk retrieval though text similarity we chose to use
the same concept of examples as for PBE
% TODO Moritz: your question was: More details on this.
% How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM.
% word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation?
and apply the Vector Space Model~\cite{schutze2008introduction}
as follows.
The
lines of the output strings of the training examples define our search
query.
The algorithm splits the search query into single lines and
identifies tokens, in our case words.
Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words.
Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

\subsubsection{Application}
To retrieve the desired information from a build log, we parse the
whole text and process it in the same way as the output of the
training examples.
The algorithm calculates the cosine
similarity~\cite{korenius2007principal} to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the outputs of the training examples determines how many of
the most similar lines are returned as the output of the retrieval
run.

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad-hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason the build
failed within a log while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.
% As this is a technique readily available in many tools developers
% use to view build logs, we study when such a keyword search is
% suitable for retrieving information chunks from CI build logs.
% In
% the following we will explain how we use simple keyword search to
% retrieve information from CI build logs, which we refer to as
% \emph{KWS}\@.

@@

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}
In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log.
RLR essentially mimicks the
situation of guessing purely blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.

@@

\subsection{Other Techniques}
% TODO: we could add more techniques from the lit.
% survey here
% TODO: probably remove
%Literature mentions further build log analysis techniques.
%This
%section describes them with our notion of chunk retrieval techniques.

\noindent
\textbf{Log Diff}
Amar et al.\ use a technique based on line diffs and information
retrieval to identify relevant lines from a failed build
log~\cite{amar2019mining}.
The configuration for the technique
is the log from the last successful build and relevant past failures.
This technique retrieves the lines from a build log that are not
present in the successful build log and contain terms related to the
given past failures.

\subsection{Tool Implementation}
For our comparison study we implemented PBE, CTS, KWS and RLR and a
unifying interface.
The unified interface is implemented in Ruby and
calls the separate technique implementations over the command line.
The implementation of PBE in C\# is based on the Microsoft PROSE
library~\cite{prose2019webpage}.
We implemented CTS, KWS and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.

@@

To investigate which technique is best suited to retrieve chunks from
CI build logs we evaluate the three chunk retrieval techniques
PBE, CTS and KWS, described in Sections~\ref{sec:expl-pbe},
\ref{sec:expl-ts} and \ref{sec:expl-skws}.
Random Line Retrieval
(RLR), explained in Section~\ref{sec:expl-rlr}, acts as a baseline for
the comparison.
This section
describes our study design and which metrics we measure to answer our
research questions.
In the presentation of the results, we first focus
on each of the three techniques and later compare them against each
other.

>>>>

% \subsubsection{Build Logs}
% often focussed on specific logs, sometimes mentioned because of effort
% developing a parser other times also because other tools used in
% further research are limiting the scope (reference paper with PIT
% tool).
% many focus on maven because of it's prevalence also in respect
% to other tools much research done on travis ci logs b/c of their
% abundance and open availability
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval
techniques we study in this article, namely program synthesis by
example (PBE), common text similarity (CTS), and keyword search (KWS).

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
In this article, we evaluate different techniques to automatically
retrieve pieces of information that appear literally in the build
logs.
%, i.e., techniques that do not aggreagte, combine, or deducted
%information.
We call such pieces of information from build logs
\emph{chunks}, and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require a formal lexer and
parser to analyze the entire structure of build logs, but focus on
ad-hoc extracting just one specific piece of information per
configuration.

@@

\begin{table*}[]
\centering
\caption{Chunk retrieval techniques.}
\begin{tabularx}{\textwidth}{@{}XlXlXX@{}}
\toprule
Name			     & Acronym & Identification Technique
& Granularity & Configuration & Source		  \\
\midrule
Program Synthesis by Example & PBE     & Regular expression program
& Character   & In/output examples	\\
Common Text Similarity	     & CTS     & TF-IDF \& cosine similarity,
expected number of lines & Line        & Output examples	   \\
Keyword Search		     & KWS     & Keywords, expected number of
lines			 & Line        & Keywords, context length  \\
Random Line Retrieval	     & RLR     & Random sample
& Line	      & Retrieval length	  \\
\bottomrule
\end{tabularx}
\label{tab:ctr}
\end{table*}

\subsection{Program Synthesis by Example (PBE)}
\label{sec:expl-pbe}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this apprach, we implemented a chunk retrieval technique
for build logs.
Our implementation is based and highly influenced by the
\emph{Microsoft PROSE library}~\cite{prose2019webpage}.
The PROSE library is powered by the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example, we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that should be retrieved by the synthesized program when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.\
because the regular expression
necessary would be too complicated to synthesize, PBE returns the
error message ``no program found''.
\Cref{lst:prose-program-simplified} presents a simplified version
of a program learned from the examples in \Cref{lst:chunk-example}.
% TODO Caro: check that these figures are latest inserted here

@@

\subsection{Common Text Similarity (CTS)}
\label{sec:expl-ts}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
We investigate a chunk retrieval technique based on the common
Vector Space Model~\cite{schutze2008introduction}.
The aim is to retrieve those lines from a build log which are most
similar to examples given by the user.

\subsubsection{Configuration}
To configure chunk retrieval though text similarity we chose to use
the same concept of examples as for PBE
% TODO Moritz: your question was: More details on this.
% How do you
% 'apply' VSM? On what?
% all the following sentences in this subsubsection describe our way
% of determining text similarity which as far as I understand and
% Annibale said is the VSM.
% word tokens, tf-idf, similarity measure
% I added a "as follows", think that is enough for explanation?
and apply the Vector Space Model~\cite{schutze2008introduction}
as follows:
The lines of the output strings of the training examples define our
search query.
The algorithm splits the search query into single lines and
identifies tokens, in our case words.
Then we build a
document-term-frequency matrix over the lines from the search query
and prune very often or very rarely appearing words.
Finally, the
algorithm applies TF-IDF to the matrix, a best practice for natural
language queries~\cite{lee1997document}.

\subsubsection{Application}
To retrieve the desired information from a build log, we parse the
whole text and process it in the same way as the search query.
The algorithm calculates the cosine
similarity~\cite{korenius2007principal} to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the outputs of the training examples determines how many of
the most similar lines are returned as the output of the retrieval
run.

\subsection{Keyword Search (KWS)}
\label{sec:expl-skws}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad-hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason a build
failed while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.

@@

\subsection{Random Line Retrieval}
\label{sec:expl-rlr}
In our evaluation, we want to compare against a baseline of randomly
picking lines from the build log.
RLR mimicks the
situation of guessing blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
return.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples.

@@

\subsection{Tool Implementation}
For our comparison study we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is implemented in Ruby and
calls the separate technique implementations over the command line.
The implementation of PBE in C\# is based on the Microsoft PROSE
library~\cite{prose2019webpage}.
We implemented CTS, KWS and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.

@@

To investigate which technique is best suited to retrieve chunks from
CI build logs we evaluate the three chunk retrieval techniques
PBE, CTS and KWS.
Random Line Retrieval (RLR), acts as a baseline for
the comparison.
The following section
describes our study design and which metrics we measure to answer our
research questions.
In the presentation of the results, we first focus
on each of the three techniques and later compare them against each
other.
